CLPFUDatabase: A suite of R packages for energy conversion chain analysis

### 2023-12-20 README.Rmd
Added metapackage information
PFUDatabase-->PFUPipeline; PFUAggDatabase-->PFUAggPipeline

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

We created a suite of `R` packages 
to support the creation and use of the CL-PFU database.
The CL-PFU database is a new resouce 
for the SEA research community and
assists SEA practitioners 
to analyze energy conversion chains (ECCs).
The new packages enable, for the first time, scalable SEA.

@@

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |

>>>>

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfupipeline]:      https://github.com/EnergyEconomyDecoupling/PFUPipeline
[link-pfuaggpipeline]:   https://github.com/EnergyEconomyDecoupling/PFUAggPipeline
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

We created a suite of `R` packages 
and the metapackage `CLPFUDatabase`
to support the creation and use of the CL-PFU database.
The CL-PFU database is a new resouce 
for the SEA research community and
assists SEA practitioners 
to analyze energy conversion chains (ECCs).
The new packages enable, for the first time, scalable SEA.

@@

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUPipeline][link-pfupipeline] and [PFUAggPipeline][link-pfuaggpipeline] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUPipeline][link-pfupipeline]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggPipeline][link-pfuaggpipeline] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |

### 2023-12-20 paper.Rmd
PFUDatabase-->PFUPipeline; PFUAggDatabase-->PFUAggPipeline

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfudatabase]:      https://github.com/EnergyEconomyDecoupling/PFUDatabase
[link-pfuaggdatabase]:   https://github.com/EnergyEconomyDecoupling/PFUAggDatabase
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-clpfudatabase]:    https://github.com/EnergyEconomyDecoupling/CLPFUDatabase
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

The steps to create the PSUT matrices for each country and each year
are accomplished by a
[targets][link-targets] [@Landau:2021aa]
computation pipeline
available in the [PFUDatabase][link-pfudatabase]
package.
A unique feature of the
[PFUDatabase][link-pfudatabase]
pipeline enabled by the [Recca][link-recca] package 
is an innovative exemplar system that allows
analyses to proceed when allocation or efficiency data are unavailable for a country.
When allocations or efficiencies are missing,
a string of exemplar countries or regions
are queried for the required information.
For example, exemplars for Belgium are
France, Europe, and ultimately, the World, in that order.
The design of the calculation pipeline in the [PFUDatabase][link-pfudatabase] package 
allows allocation and efficiency data
for any country or year to be added at a later date 
to improve the database.

A second [targets][link-targets] pipeline 
in the [PFUAggDatabase][link-pfuaggdatabase] package [@Heun-PFUAggDatabase:2023]
aggregates ECCs
(a) by region 
(e.g., continents and world),
(b) by energy carrier (product) category 
(e.g., Coal and coal products, 
Low-, Medium-, and High-temperature heat, etc.), 
(c) by energy conversion machines (industries) and/or end use sectors
(e.g., Residential, Transport, etc.), and
(d) to primary, final, and useful (PFU) stages.
In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
calculates efficiencies for each country
at all available aggregations.
The [PFUAggDatabase][link-pfuaggdatabase]
pipeline uses the [Recca][link-recca] package extensively.

Both [targets][link-targets] pipelines 
(in the [PFUDatabase][link-pfudatabase] and
[PFUAggDatabase][link-pfuaggdatabase] packages)
produce data frames readable by the [pins][link-pins] package [@Silga:2023].
Both pipelines benefit from the 
[PFUSetup][link-pfusetup] [@Heun-PFUSetup:2023] and
[PFUPipelineTools][link-pfupipelinetools] [@Heun-PFUPipelineTools:2023]
packages. 
[PFUSetup][link-pfusetup] identifies storage locations 
for pipeline input and output data.
[PFUPipelineTools][link-pfupipelinetools] provides functions and constants
common to both pipelines.

@@

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUDatabase][link-pfudatabase] and [PFUAggDatabase][link-pfuaggdatabase] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUDatabase][link-pfudatabase]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggDatabase][link-pfuaggdatabase] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all package dependencies for the CL-PFU database |

@@

# (3) Run the PFUDatabase pipeline to create RUVY matrices.
targets::tar_make_future(
  script = file.path("ExampleFolder", "_targets_pfudatabase.R"), 
  workers = 4
)

# (4) Run the PFUAggDatabase pipeline to aggregate 
#     energy and calculate efficiencies.
targets::tar_make_future(
  script = file.path("ExampleFolder", "_targets_pfuaggdatabase.R"),
  workers = 4
)
```

```{r}
# (5) Review results using the pins package.
#     Establish the pinboard.
pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
  pins::board_folder(versioned = TRUE)
#     Look at the psut data frame
#     created by the PFUDatabase pipeline.
psut <- pinboard |>
  pins::pin_read(name = "psut", version = "20231219T204403Z-2c3ce")
head(as.data.frame(psut), 2)
#     Each column of the psut data frame contains sparse matrices, 
#     the columns of which are suppressed on display.
#     Show the column names here.
colnames(psut$R[[1]])
#     Show an example resource (R) matrix.
psut$R[[1]]

#     Review the aggregate PFU efficiency data frame
#     created by the PFUAggDatabase pipeline.
agg_eta_pfu <- pinboard |> 
  pins::pin_read(name = "agg_eta_pfu", version = "20231211T161827Z-8f5e7")
agg_eta_pfu |> 
#     Filter and delete columns 
#     to present a subset of exergy results in a data frame.
#     The EX.p, EX.f, and EX.u columns contain aggregate exergy at
#     primary, final, and useful stages, respectively, in TJ.
#     The eta_pf, eta_fu, and eta_pu columns contain aggregate
#     primary-to-final, final-to-useful, and 
#     primary-to-useful efficiencies.
  dplyr::filter(Energy.type == "X", IEAMW == "Both", 
                Product.aggregation == "Specified", 
                Industry.aggregation == "Specified", 
                GrossNet == "Net",
                Country %in% c("GHA", "ZAF")) |> 
  dplyr::mutate(
    Method = NULL, 
    Energy.type = NULL,
    Product.aggregation = NULL,
    Industry.aggregation = NULL,
    IEAMW = NULL,
    Chopped.mat = NULL, 
    Chopped.var = NULL, 
    GrossNet = NULL
  )
```

>>>>

<!-- Define some macros for later use -->
[link-cran]:             https://cran.r-project.org
[link-ecctools]:         https://github.com/earamendia/ECCTools
[link-eroitools]:        https://github.com/earamendia/EROITools
[link-iea]:              https://www.iea.org
[link-ieatools]:         https://github.com/MatthewHeun/IEATools
[link-ieaeweb]:          https://www.iea.org/data-and-statistics/data-product/world-energy-balances
[link-ilo]:              https://www.ilo.org/
[link-fao]:              https://www.fao.org/
[link-matsbyname]:       https://github.com/MatthewHeun/matsbyname
[link-matsindf]:         https://github.com/MatthewHeun/matsindf
[link-mwtools]:          https://github.com/EnergyEconomyDecoupling/MWTools
[link-pfusetup]:         https://github.com/EnergyEconomyDecoupling/PFUSetup
[link-pfupipeline]:      https://github.com/EnergyEconomyDecoupling/PFUPipeline
[link-pfuaggpipeline]:   https://github.com/EnergyEconomyDecoupling/PFUAggPipeline
[link-pfupipelinetools]: https://github.com/EnergyEconomyDecoupling/PFUPipelineTools
[link-clpfudatabase]:    https://github.com/EnergyEconomyDecoupling/CLPFUDatabase
[link-pins]:             https://pins.rstudio.com
[link-rclabels]:         https://github.com/MatthewHeun/RCLabels
[link-recca]:            https://github.com/MatthewHeun/Recca
[link-targets]:          https://docs.ropensci.org/targets
[link-tidyverse]:        https://www.tidyverse.org

@@

The steps to create the PSUT matrices for each country and each year
are accomplished by a
[targets][link-targets] [@Landau:2021aa]
computation pipeline
available in the [PFUPipeline][link-pfupipeline]
package.
A unique feature of the
[PFUPipeline][link-pfupipeline]
pipeline enabled by the [Recca][link-recca] package 
is an innovative exemplar system that allows
analyses to proceed when allocation or efficiency data are unavailable for a country.
When allocations or efficiencies are missing,
a string of exemplar countries or regions
are queried for the required information.
For example, exemplars for Belgium are
France, Europe, and ultimately, the World, in that order.
The design of the calculation pipeline in the [PFUPipeline][link-pfupipeline] package 
allows allocation and efficiency data
for any country or year to be added at a later date 
to improve the database.

A second [targets][link-targets] pipeline 
in the [PFUAggPipeline][link-pfuaggpipeline] package [@Heun-PFUAggPipeline:2023]
aggregates ECCs
(a) by region 
(e.g., continents and world),
(b) by energy carrier (product) category 
(e.g., Coal and coal products, 
Low-, Medium-, and High-temperature heat, etc.), 
(c) by energy conversion machines (industries) and/or end use sectors
(e.g., Residential, Transport, etc.), and
(d) to primary, final, and useful (PFU) stages.
In addition, the [PFUAggDatabase][link-pfuaggdatabase] pipeline 
calculates efficiencies for each country
at all available aggregations.
The [PFUAggDatabase][link-pfuaggdatabase]
pipeline uses the [Recca][link-recca] package extensively.

Both [targets][link-targets] pipelines 
(in the [PFUPipeline][link-pfupipeline] and
[PFUAggPipeline][link-pfuaggpipeline] packages)
produce data frames readable by the [pins][link-pins] package [@Silga:2023].
Both pipelines benefit from the 
[PFUSetup][link-pfusetup] [@Heun-PFUSetup:2023] and
[PFUPipelineTools][link-pfupipelinetools] [@Heun-PFUPipelineTools:2023]
packages. 
[PFUSetup][link-pfusetup] identifies storage locations 
for pipeline input and output data.
[PFUPipelineTools][link-pfupipelinetools] provides functions and constants
common to both pipelines.

@@

| Package | Purpose |
|:--------|:--------|
| [IEATools][link-ieatools]             | Converts [IEA EWEB][link-ieaeweb] data to RUVY matrices in [matsindf][link-matsindf] data frames |
| [MWTools][link-mwtools]               | Converts [ILO][link-ilo] and [FAO][link-fao] data to RUVY matrices of human and animal muscle work in [matsindf][link-matsindf] data frames |
| [Recca][link-recca]                   | Performs `R` energy conversion chain analysis |
| [PFUSetup][link-pfusetup]             | Identifies input and output locations for the [PFUPipeline][link-pfupipeline] and [PFUAggPipeline][link-pfuaggpipeline] pipelines |
| [PFUPipelineTools][link-pfupipelinetools] | Provides basic functionality for all PFU pipelines |
| [PFUPipeline][link-pfupipeline]       | Provides a [targets][link-targets] pipeline to create a data frame of RUVY matrices |
| [PFUAggPipeline][link-pfuaggpipeline] | Provides a [targets][link-targets] pipeline to aggregate RUVY matrices |
| [CLPFUDatabase][link-clpfudatabase]   | A metapackage (like `tidyverse`) to quickly load all package dependencies for the CL-PFU database |

@@

# (3) Run the PFUPipeline to create RUVY matrices.
targets::tar_make_future(
  script = file.path("ExampleFolder", "_targets_pfupipeline.R"), 
  workers = 4
)

# (4) Run the PFUAggPipeline to aggregate 
#     energy and calculate efficiencies.
targets::tar_make_future(
  script = file.path("ExampleFolder", "_targets_pfuaggpipeline.R"),
  workers = 4
)
```

```{r}
# (5) Review results using the pins package.
#     Establish the pinboard.
pinboard <- file.path("ExampleFolder", "OutputData", "PipelineReleases") |> 
  pins::board_folder(versioned = TRUE)
#     Look at the psut data frame
#     created by the PFUPipeline.
psut <- pinboard |>
  pins::pin_read(name = "psut", version = "20231219T204403Z-2c3ce")
head(as.data.frame(psut), 2)
#     Each column of the psut data frame contains sparse matrices, 
#     the columns of which are suppressed on display.
#     Show the column names here.
colnames(psut$R[[1]])
#     Show an example resource (R) matrix.
psut$R[[1]]

#     Review the aggregate PFU efficiency data frame
#     created by the PFUAggPipeline.
agg_eta_pfu <- pinboard |> 
  pins::pin_read(name = "agg_eta_pfu", version = "20231211T161827Z-8f5e7")
agg_eta_pfu |> 
#     Filter and delete columns 
#     to present a subset of exergy results in a data frame.
#     The EX.p, EX.f, and EX.u columns contain aggregate exergy at
#     primary, final, and useful stages, respectively, in TJ.
#     The eta_pf, eta_fu, and eta_pu columns contain aggregate
#     primary-to-final, final-to-useful, and 
#     primary-to-useful efficiencies.
  dplyr::filter(Energy.type == "X", IEAMW == "Both", 
                Product.aggregation == "Specified", 
                Industry.aggregation == "Specified", 
                GrossNet == "Net",
                Country %in% c("GHA", "ZAF")) |> 
  dplyr::mutate(
    Method = NULL, 
    Energy.type = NULL,
    Product.aggregation = NULL,
    Industry.aggregation = NULL,
    IEAMW = NULL,
    Chopped.mat = NULL, 
    Chopped.var = NULL, 
    GrossNet = NULL
  )
```
